---
title: "1.Tern_wind_visualization"
author: "Dennis Kim"
date: '2022-05-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preamble 

call relevant libraries 
```{r import_libraries}
# data wrangling 
library(dplyr)
library(readr)
library(data.table)
library(ezknitr)
library(here)
library(DT)
library(lubridate)
library(stringr)
library(bit64)
library(tidyr)
library(lubridate)
library(adehabitatLT)

# visualization
library(moveVis)
library(move)
library(ggplot2)
library(gganimate)
library(magick)
library(ggOceanMaps)
library(ggspatial)
library(grid)
library(transformr)
library(gifski)
library(magick)
library(cmocean)

# environmental layers 
library(raster)
library(rWind)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
```

# Environmental data

Import the raw tracking data from your designated folder 

## world map data 
```{r ocean data}
# call the sf object of world map from rnaturalearth pacakge
world <- ne_countries(scale = "medium", returnclass = "sf")
```

## daily wind data
```{r wind data}
# call the wind_data rds
wind <- read_rds(here("data", "modified_fall_wind_daily.Rdata"))

# summary of the filtered wind data
summary(wind)
```

# Movement data

## tern data

Import the tracking data (BF426)
```{r tern data}
# tern data

## call the raw filtered csv data - name it as its id 
BF426 <- fread(here("data/ModeledTracks/BF426_smNEW.csv"))

## create a new column with id 
BF426 <- BF426 %>% mutate(names = "BF426")

## only select fall migration (2017-08-08 12:22:00 ~ 2017-11-17 12:22:00)
fall.BF426 <- BF426 %>% filter(Time1 > '2017-08-08 00:00:00' & Time1 < "2017-11-17 23:59:59")

## summary of terns 
summary(fall.BF426)
```

Find the regularized time stamps for the individuals 
```{r fall migratiom BF426}
# Table of time intervals in data 
plot(table(diff(fall.BF426$Time1)), xlim = c(0,24), xlab = "time interval (hour)", ylab = "count")
```
It looks like 11 hours might be the best regularized timestamp for this individual. 

Regular time intervals using NAs 
```{r regularized the time}
# select the columns that we want only
trk <- fall.BF426 %>% dplyr::select(Time1, `Lon.50%`, `Lat.50%`, names)

# change the column names 
colnames(trk) <- c("time","x","y","ID")

# convert the data to move object
fall.tern <- move(x = trk$x, y = trk$y, time = as.POSIXct(trk$time, format = "%Y-%m-%d %H:%M:%OS", tz="UTC"),
                  data = trk, animal= trk$ID)

# align the movement data to a uniform time scale (11 hrs)
fall.tern.mv <- align_move(fall.tern, res = 11, unit = "hours", spaceMethod = "greatcircle")

# check if it was correctly regularized 
summary(fall.tern.mv)

# convert the moveobject back to data frame 
mvDF <- as.data.frame(fall.tern.mv)

# change the column to id 
colnames(mvDF)[10] <- "id"

# convert data time as integer 
mvDF$time = as.Date(mvDF$time)

# check
head(mvDF)

# save the move object of BF426
#write_rds(mvDF, path = here("data", "mvDF_daily.Rdata"))
```

Since the tracking points are multiple within a day, we would like to average them out per day so that it matches with the wind time periods and show one point per day when we simulate
```{r further trk modification}
# avg location per day 
avg.mvDF = mvDF %>% group_by(time, id) %>% 
  summarise(x = mean(x),
            y = mean(y))

# check the data 
summary(avg.mvDF)

# save the avg move object of BF426
#write_rds(avg.mvDF, path = here("data", "avg_mvDF_daily.Rdata"))
```


## Visualization
```{r mv and wind visualization}
# visualization 
tern.move.wind.vector = ggplot()+
  
  # wind layer
  geom_raster(data = wind, aes(x = lon, y = lat, fill = velocity))+
  geom_segment(data = wind,
               aes(x = lon, xend = lon+u/10, y = lat, yend = lat+v/10), arrow = arrow(length = unit(0.2, "cm")))+
  scale_fill_gradientn(colours = cmocean('speed')(320), limits = c(0,32), 
                       na.value = "white", name = "Speed\n (m/s)")+
  
  # environmental layer
  geom_sf(data = world, fill = "grey85", col = 1)+
  coord_sf(xlim = c(-80, 30), ylim = c(-70, 70), expand = FALSE)+
  
  # movement track 
  geom_point(data = avg.mvDF, aes(x = x, y = y, colour = "BF426"), fill = "red",alpha = 1, shape = 21, size = 4) + 
  #geom_path(data = avg.mvDF, aes(x = x, y = y, color = id), size = 2) +
  scale_color_manual(values = c("red"), name = "ID")+
  
  # formatting
  scale_x_continuous(breaks = c(-80,30))+
  theme_bw()

# check the ggplot visualization
#tern.move.wind.vector

# gganimate specifics 
tern.anim = tern.move.wind.vector+ transition_time(time) +
  labs(x = NULL, y = NULL, title = "Arctic Tern BF426: {frame_time}")+ 
  ease_aes("linear")

# animate the visualization
animate(tern.anim, height = 1000, width = 1000, fps = 5)

# save the animation 
anim_save("fall_BF426_wind.gif", width = 1000, height = 1000)
```


# Document Footer
```{r}
sessionInfo()
```