---
title: "2.data-prep.Rmd"
author: "Dennis Kim"
date: "5/27/2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preamble 

call relevant libraries 
```{r import_libraries}
# data wrangling 
library(dplyr)
library(readr)
library(data.table)
library(ezknitr)
library(here)
library(DT)
library(lubridate)
library(stringr)
library(bit64)
library(tidyr)
library(purrr)

# environmental layers 
library(raster)
library(sf)
library(rWind)
library(stars)
library(gdistance)

# movement analysis 
library(amt)
```

# Import Dataset 

## tern movement data

Import the tracking data (BF426)
```{r tern data}
# call the raw filtered csv data - name it as its id 
BF426 <- fread(here("data/ModeledTracks/BF426_smNEW.csv"))

# create a new column with id 
BF426 <- BF426 %>% mutate(names = "BF426")

# only filter fall migration (2017-08-08 12:22:00 ~ 2017-11-17 12:22:00)
fall.BF426 <- BF426 %>% filter(Time1 > '2017-08-08 00:00:00' & Time1 < "2017-11-17 23:59:59")

# select the columns that we want only
fall.BF426 <- fall.BF426 %>% dplyr::select(Time1, `Lon.50%`, `Lat.50%`, names)

# change the column names 
colnames(fall.BF426) <- c("time","x","y", "id")

# create an additional column the data format without time 
fall.BF426$date = as.Date(fall.BF426$time)

# summary of terns 
summary(fall.BF426)
```

## daily wind data
```{r wind data}
## set the time series 11 hrs interval
#fall.wind <- seq(ymd_hms(paste(2017, 8, 8, 00, 00, 00, sep ="-")), ymd_hms(paste(2017, 11, 17, 23, 59, 59, sep = "-")), by = "24 hours")

# download Atlantic ocean dataset of a specified date range from GFS 
#wind_data <- wind.dl_2(fall.wind, -80, 30, -70, 70)

# tidy the wind data as a data frame
#fall_wind <- rWind::tidy(wind_data)

# save the wind_data 
#write_rds(fall_wind, path = here("data", "fall_wind_daily_unfiltered.Rdata"))

# read the rds file (the saved above code runs)
fall.wind <- read_rds(here("data", "fall_wind_daily_unfiltered.Rdata"))

# look at the summary of the data
summary(fall.wind)

# convert the wind data to raster
wind_raster = wind2raster(fall.wind)

# plot the raster
plot(wind_raster)

# calculate flow-based cost based on paper Fellcislmo et al. 2008
## cost = (1/speed)*(horizontal factor)
wind_cost <- flow.dispersion(wind_raster, type = "active")

# check the cost transition object
wind_cost

# convert it to the raster object 
wind_cost_raster =  raster(wind_cost)

# plot the raster object 
plot(wind_cost_raster)
```

# data manipulation 

## amt object: make tracks and extract wind layer values per track

convert the tern tracking data to trk object (resample by 12 hours and create available steps)
```{r convert trk object}
# create a track of BF426
trk.BF426 <- amt::make_track(fall.BF426, .x = x, .y = y, .t = time, id = id, date = date)

# check the sampling rates of the data - 12 hours might be the best
trk.BF426 %>% summarize_sampling_rate()

# re-sample the tracks by 12 hours and create used and available points and steps 
resample_trk_BF426 <- trk.BF426 %>% track_resample(rate = hours(12), tolerance = hours(1)) %>% steps_by_burst() %>% random_steps()
```

extract wind layer values 
```{r extract wind speed and dir values}
# extract fall wind direction, speed, and layer
fall_wind_trk_BF426 <- resample_trk_BF426 %>% extract_covariates(wind_raster) %>% extract_covariates(wind_cost_raster)

# check the data if the values are successfully extracted from the layers 
summary(fall_wind_trk_BF426)

# change the wind cost column name as "cost"
colnames(fall_wind_trk_BF426)[15] <- "cost"
```

save the extracted resampled trk data 
```{r save the resampled trk data}
# save the filtered saved trk data with wind covariates
#write_rds(fall_wind_trk_BF426, path = here("data", "fall_wind_trk.Rdata"))
```

# Document Footer
```{r}
sessionInfo()
```