---
title: "3.data_analysis"
author: "Dennis Kim"
date: '2022-07-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preamble 

call relevant libraries 
```{r import_libraries}
# data wrangling 
library(dplyr)
library(readr)
library(data.table)
library(ezknitr)
library(here)
library(DT)
library(lubridate)
library(stringr)
library(bit64)
library(tidyr)
library(purrr)
library(tidyr)
library(broom)
library(forcats)

# movement analysis 
library(amt)

# visualization
library(ggplot2)
library(jtools)
library(ggstance)
library(broom.mixed)
```

# Import Dataset 

Read in the tern tracks 
```{r tern tracks}
# read the tern tracks 
ssfdat <- read_rds(here("data", "fall_wind_trk.Rdata"))

# summary of the data 
summary(ssfdat)
```

# Fit SSF
```{r fit ssf}
# model fit
m1 <- ssfdat %>% fit_issf(case_ ~ direction + speed + cost + sl_ + log(sl_) + cos(ta_)+ strata(step_id_), model = TRUE)

# summary of coefficients
summary(m1)
```
**[Interpreation]**

* direction: **negative coefficient: the tern is less likely to select the location as the direction increases by 1 unit while others are equally available.**
* speed: **negative coefficient: the tern is less likely to select the location as the speed increases by 1 unit while others are equally available.**
* cost: **negative coefficient: the tern is less likely to select the location as the cost increases by 1 unit while others are equally available.**

# Visualization 

## Prep for coefficient data 
```{r coefficient df}
# coefficient data frame 
coef_df <- m1$model$coefficients %>% tidy()

# CI data frame
conf_df <- m1$model %>% confint() %>% as.data.frame() %>% mutate(names = c("direction", "speed", "cost", "sl_", "log(sl_)", "cos(ta_)"))

# left join the coef df to conf df 
fit.df <- left_join(conf_df, coef_df)

# change the column names 
colnames(fit.df) <- c("lowCI", "highCI", "names", "beta")

# reorder by column name 
fit.df <- fit.df[, c("names", "beta", "lowCI", "highCI")]

# check the filtered data 
fit.df
```

## Visualization of the coefficients with CIs 
```{r visualization with CIs}
# visualization 
tern.fig <- fit.df %>% dplyr::mutate(names = forcats::fct_relevel(names, sort)) %>% 
  ggplot(., aes(beta, names, colour = names)) +
  geom_vline(xintercept = 0, lty = 2, lwd = 1, colour = "grey50")+
  geom_point(size = 3, pch = 20, position = position_dodge(width = 0.5))+
  geom_linerange(aes(xmax = highCI, xmin = lowCI))+
  theme_bw()+
  xlab("Effect size")+
  ggtitle("SSF Parameter estimates")+
  theme(legend.position = "none",
        axis.title.y = element_blank())

tern.fig
```

save the SSF estimates plot 
```{r save the plot}
ggsave("tern_fig.png", tern.fig, width = 10, height = 7)
```


# Document Footer
```{r}
sessionInfo()
```